---
title: "Read in NSW Cancer Council Data"
output: html_document
---

For this data reading, I have used SQL commands and information provided by Maddi to extract data from the docker image into CSV files

```{r setup}
# Copyright 2018 James Pearce, 3 Crowns Consulting Pty Ltd
# library(dplyr)
library(magrittr)
library(data.table)
# library(dtplyr)
```



```{r}

data_path <- "./data/NSW Cancer Council"
data_files <- c("communication_history.csv",
                "dim_campaign.csv"
                )
```

Because SQL Server `sqlcmd` spits out data in a semi-formatted way (and that's what I used), the second line needs to be removed.  Create a function for reading in the data.

```{r}
fread_sqlserver <- function(file){
  dt_body <- fread(file, skip=2, na.strings="NULL")
  dt_head <- fread(file, nrows=0)
  setnames(dt_body, names(dt_head))
  return(dt_body)
}
```


```{r}
campaign_DT <- fread_sqlserver(file.path(data_path, data_files[2]))
```

```{r}
comms_history_DT <- fread_sqlserver(file.path(data_path, data_files[1]))
```

Save the files in a better format.

```{r}
saveRDS(comms_history_DT, "comms_history_dt.rds")
```
```{r}
saveRDS(campaign_DT, "campaign_dt.rds")
```

### Constituent information

`dbo.ConstituentView`
```{r}
constituent_view_dt <- fread_sqlserver(file.path(data_path, 
                                                 "constituent_view.csv"))
```

`Dim.Constituent` has not been read in. It doesn't contain segments and has a lot of fields, requiring the maximum row to be greater than 1024.

Though you would have to think `dbo.ConstituentRFMSegment` will contain RFM segment information. Let's examine.

```{r}
constituent_RFM_dt <- fread_sqlserver(file.path(data_path, 
                                                 "constituent_RFM.csv"))
```
```{r}
saveRDS(constituent_RFM_dt, "constituent_RFM.rds")
```

```{r}
donations_dt <- fread_sqlserver(file.path(data_path, "donations.csv"))
```


```{r}
saveRDS(donations_dt, file.path(data_path, "donations.rds"))
```

```{r}
# constituent_DT <- fread(file.path(data_path, data_files[2]))
# didn't work
# campaign_DT <- fread(file.path(data_path, data_files[7]))
# didn't work
constituent_participation_history_DT <- fread(file.path(data_path,
                                                        data_files[3]))
```
```{r}
saveRDS(constituent_participation_history_DT, file=file.path(data_path, "constituent_participation_history.rds"))
```

```{r}
constituent_summary_DT <- fread(file.path(data_path, data_files[5]))
```

```{r}
saveRDS(constituent_summary_DT, file.path(data_path, "constituent_summary.rds"))
```

```{r}
constituent_communication_preferences_DT <- fread(file.path(data_path,
                                                            data_files[1]))
```

```{r}
constituent_communication_preferences_DT %>% head(10)
```

Copyright 2018 James Pearce, 3 Crowns Consulting Pty Ltd
