---
title: "03 Explore data"
output: html_notebook
---

Explore the NSW Cancer Council data.

```{r setup}
library(data.table)
library(magrittr)
library(lubridate)
```


Read in the data from the RDS.



```{r}
comms_history_dt <- readRDS("comms_history_dt.rds")
donations_dt <- readRDS('data/NSW Cancer Council/donations.rds')
campaign_dt <- readRDS("campaign_dt.rds")

```


### Create cut-down versions
```{r}
comms_history_sample <- comms_history_dt[(ProgramName == "Appeals / OG") & (DateSent >= "2015-04-01"), ]
saveRDS(comms_history_sample, "comms_history_sample.rds")
```

```{r}
comms_history_sample <- readRDS("comms_history_sample.rds")
```

```{r}
comms_history_sample[year(DateSent) == 2017, .N, by=.(CampaignGroupName, AppealCode)][order(CampaignGroupName, -N)]
```
```{r}
comms_history_sample[year(DateSent) == 2017] %>% head
```
```{r}
channel_lookup <- unique(campaign_dt, by=c("AppealCode", "ChannelName"))[ChannelName != "Not Applicable" & AppealCode != "Not Applicable", 
                                                       .(AppealCode, ChannelName)]
channel_lookup[, .N, by=.(AppealCode)][N > 1]
comms_history_sample[year(DateSent) == 2017, .N]
comms_history_sample <- channel_lookup[comms_history_sample, on=.(AppealCode)] 
```



```{r}
setorder(comms_history_sample, ConstituentSKey, DateSent)
```

Look at the channels by appeal code.
```{r}
comms_history_sample[year(DateSent) == 2017, .N, by=.(AppealCode, ChannelName)][order(-N)]
```


Keen only the responses whose constituents match with the communications history.
```{r}
constits <- comms_history_sample[, ConstituentSKey] %>% unique
donations_sample <- donations_dt[DonorConstituentSKey %in% constits,]
setorder(donations_sample, DonorConstituentSKey)
saveRDS(donations_sample, "donations_sample.rds")
```

```{r}
# comms_history_unique <- comms_history_sample[, .(ConstituentSKey, ProgramName, CampaignGroupName, CampaignUnitName,
#                                                  CampaignYear, AppealCode, PackageCode_clean, PackageName_clean)] %>% 
#   unique 
```

Merge the top 3 appeals with the donations table.

```{r}
donation_responses <- merge(comms_history_sample,
      donations_sample, 
      by.x=c("ConstituentSKey", "PackageCode_clean"),
      by.y=c("DonorConstituentSKey", "PackageID"),
      all.x=TRUE
      )
saveRDS(donation_responses, "donation_responses.rds")
```



```{r}
# Clean up the stuff
rm(comms_history_dt, donations_dt)
gc()
```

## Create the features
```{r}
date_this <- '2017-04-01' %>% as.Date
recency_ft <- function(date, date_current, duration_last=ddays(365)){
  # date : vector of dates to transform
  # date_current : the assumed current date (ignore after this)
  # duration_last : earliest date to consider as measured from date_current (duration)
  # **return** : days since last transaction

  
  answer <- as.duration(date_current - date) %>%
    as.numeric("days")
  answer[is.na(date)] <- NA
  answer[date > date_current] <- NA
  answer[date < (date_current - duration_last)] <- duration_last %>% as.numeric("days")
  return(answer)
} 

dates <- c("2017-02-04", "2017-01-01", "2017-04-02", "2015-04-02", "2017-02-05", NA) %>% as.Date
recency_ft(dates, date_this)


```
```{r}
date_this <- '2017-04-01' %>% as.Date
aggregate_ft <- function(value, date, date_current, FUN=sum, duration_last=ddays(365)){
  # value : value to be aggregated
  # date : date to transform
  # date_current : the assumed current date (ignore after this)
  # FUN : aggregation function
  # duration_last : earliest date to consider as measured from date_current (duration)
  # **return** : days since last transaction
  value[date > date_current] <- NA
  value[date < (date_current - duration_last)] <- NA
  value[is.na(date)] <- NA
  answer <- FUN(value[!is.na(value)])
  return(answer)
}

dates <- c("2017-02-04", "2017-01-01", "2017-04-02", "2015-04-02", "2017-02-05", NA) %>% as.Date
values <- c(2, 4, 8, 16, NA, 12)
aggregate_ft(values, dates, date_this) # 6
aggregate_ft(values, dates, date_this, length) # 2

```
```{r}
setkey(donations_sample, DonorConstituentSKey)
donations_sample[, DonationDate := as.Date(DonationDate)]

```

```{r}
date_this <- '2017-04-01' %>% as.Date
```


```{r}
# Use this to develop features before running them on the big data set.
donations_prototype <- donations_sample[DonorConstituentSKey > -2][1:10000]
```

```{r}
min_NA <- function(vec){
  as.numeric(ifelse(all(is.na(vec)), NA, min(vec, na.rm=TRUE)))
}

min_NA(c(NA, NA, NA))
min_NA(c(1, 2, NA))
```

Takes a long time to run. Prototype the other features on a sample as well as this feature.
```{r}
run_it <- FALSE
if (run_it){
  st <- Sys.time()
  donations_features <- donations_sample[, 
                         .(recency_3y=min_NA(recency_ft(DonationDate, date_this, dyears(3))),
                          frequency_3m=aggregate_ft(DonationAmount, DonationDate, date_this, length,
                                                    ddays(90)),
                          frequency_6m=aggregate_ft(DonationAmount, DonationDate, date_this, length, ddays(180)),
                          frequency_1y=aggregate_ft(DonationAmount, DonationDate, date_this, length, dyears(1)),
                          frequency_2y=aggregate_ft(DonationAmount, DonationDate, date_this, length, dyears(2)),
                          frequency_3y=aggregate_ft(DonationAmount, DonationDate, date_this, length, dyears(3)),
                          value_3m=aggregate_ft(DonationAmount, DonationDate, date_this, sum, ddays(90)),
                          value_6m=aggregate_ft(DonationAmount, DonationDate, date_this, sum, ddays(180)),
                          value_1y=aggregate_ft(DonationAmount, DonationDate, date_this, sum, dyears(1)),
                          value_2y=aggregate_ft(DonationAmount, DonationDate, date_this, sum, dyears(2)),
                          value_3y=aggregate_ft(DonationAmount, DonationDate, date_this, sum, dyears(3))
                          ),
                            by=.(DonorConstituentSKey)]
  Sys.time() - st
}
```

Save this as it takes 1 1/2 hours.
```{r}
saveRDS(donations_features, "donations_features.rds")
```


## Communications features

Develop the following features:

  * recency and frequency as per donations
  * number of campaign groups
  * (optional) recency by campaign group
  * last communication group

Take a sample for prototyping.
```{r}
comms_history_prototype <- comms_history_sample[ConstituentSKey > -2][1:100000]
```

Need a new function to find the most recent item.

```{r}
nth_ft <- function(value, date, date_current, index=1, duration_last=ddays(365)){
  # value : value to be find the nth most recent
  # date : date to transform and sort
  # date_current : the assumed current date (ignore after this)
  # index : the nth last version to return, ie 1 = last
  # duration_last : earliest date to consider as measured from date_current (duration)
  # **return** : sorted value at index
  value[date > date_current] <- NA
  value[date < (date_current - duration_last)] <- NA
  value[is.na(date)] <- NA
  value <- value[rev(order(date))]
  value <- value[!is.na(value)]
  answer <- value[index]
  return(answer)
}

dates <- c("2017-02-04", "2017-01-01", "2017-02-02", "2015-04-02", "2017-02-05", NA) %>% as.Date
values <- c(2, 4, 8, 16, NA, 12)
nth_ft(values, dates, date_this) # 2
nth_ft(values, dates, date_this, 2) # 8

```
## Channels in the last 3 months
```{r}
comms_history_sample[, .N, by=.(ChannelName)]
```


```{r}
channel_flags <- comms_history_sample[(DateSent > (date_this - dyears(1))) & (DateSent <= date_this), length(ConstituentSKey), by=.(ConstituentSKey, ChannelName)][!is.na(ChannelName)] %>% 
  dcast(ConstituentSKey ~ ChannelName, value.var="V1")

for (j in seq_len(ncol(channel_flags)))
    set(channel_flags, which(is.na(channel_flags[[j]])), j, 0)

saveRDS(channel_flags, "channel_flags.rds")
```


```{r}
run_it <- FALSE
# run_it <- TRUE

if (run_it){
  # comms_history_features <- comms_history_prototype[, 
  comms_history_features <- comms_history_sample[,
                                                  .(
                          comm_recency_3y=min_NA(recency_ft(DateSent, date_this, dyears(3))),
                          comm_frequency_3m=aggregate_ft(PackageCode_clean, DateSent, date_this, length,
                                                    ddays(90)),
                          comm_frequency_6m=aggregate_ft(PackageCode_clean, DateSent, date_this, length, ddays(180)),
                          comm_frequency_1y=aggregate_ft(PackageCode_clean, DateSent, date_this, length, dyears(1)),
                          comm_frequency_2y=aggregate_ft(PackageCode_clean, DateSent, date_this, length, dyears(2)),
                          comm_frequency_3y=aggregate_ft(PackageCode_clean, DateSent, date_this, length,
                                                         dyears(3)),
                          comm_unique_3m=aggregate_ft(CampaignGroupName, DateSent, date_this,
                                                   function(x) length(unique(x)), ddays(90)),
                          comm_unique_6m=aggregate_ft(CampaignGroupName, DateSent, date_this,
                                                   function(x) length(unique(x)), ddays(180)),
                          comm_unique_1y=aggregate_ft(CampaignGroupName, DateSent, date_this,
                                                   function(x) length(unique(x)), dyears(1)),
                          comm_unique_2y=aggregate_ft(CampaignGroupName, DateSent, date_this,
                                                   function(x) length(unique(x)), dyears(2)),
                          comm_unique_3y=aggregate_ft(CampaignGroupName, DateSent, date_this,
                                                   function(x) length(unique(x)), dyears(3)),
                          last_comm=nth_ft(CampaignGroupName, DateSent, date_this, 1, dyears(3))
                                                  ),
                          by=.(ConstituentSKey)
                          ]
}

```

```{r}
# saveRDS(comms_history_features, "comms_history_features.RDS")
comms_history_features <- readRDS("comms_history_features.RDS")

```

Find the campaigns in the last three months before `date_this`.
```{r}
comms_history_campaign_features <- comms_history_sample[(DateSent <= date_this) & (DateSent >= (date_this - ddays(180))), 
                     .(flag=1), 
                     by=.(ConstituentSKey, CampaignGroupName)
                     ] %>%
  dcast(ConstituentSKey ~ CampaignGroupName, value.var="flag", fill=0)
```

Find the responses by `ConstituentSKey`.
```{r}
donations_response_features <- donations_sample[(DonationDate >= date_this) & (DonationDate <= (date_this + ddays(90))), 
                                               .(response=sum(DonationAmount, na.rm=T)), 
                                               by=.(ConstituentSKey=DonorConstituentSKey)]
```


Merge everything together.
```{r}
model_dt <- merge(comms_history_features, donations_features, by.x="ConstituentSKey", by.y="DonorConstituentSKey", all.x=TRUE) %>%
  merge(comms_history_campaign_features, by="ConstituentSKey", all.x=TRUE) %>%
  merge(donations_response_features, by="ConstituentSKey", all.x=TRUE)
model_dt <- model_dt[ConstituentSKey > 0]
model_dt %>% head
```
```{r}
model_dt <- merge(model_dt, channel_flags, by="ConstituentSKey", all.x=TRUE)
```

Probably need to check for NAs and change accordingly.

```{r}
model_dt <- model_dt[!is.na(comm_recency_3y)]
# get rid of the NAs

to_replace <- names(model_dt[25:32])
for (j in to_replace)
  set(model_dt, which(is.na(model_dt[[j]])), j, 0)

model_dt %>% head
```


```{r}
saveRDS(model_dt, "model_dt.rds")
```

```{r}
model_dt <- readRDS("model_dt.rds")
```

