---
title: "04.1-build-additional-logistic-model"
output: html_notebook
---

# Introduction

The purpose of this model is to create an additive component for the channel flags. We can use this in the optimisation to alter the value of the expected return.

```{r setup}
library(data.table)
library(magrittr)
library(h2o)
```

Read in the data.

```{r}
model_dt <- readRDS("model_dt.rds")
sample_prop = 0.1
set.seed(730)
model_sample <- model_dt[sample(.N, .N*sample_prop)]
```

Start up H2O.
```{r}
h2o.init()
```

Upload the data to an H2O frame.
```{r}
model_h2o <- as.h2o(model_sample, id="model_h2o")
# model_h2o$response <- as.factor(ifelse(model_h2o$response > 0, 1, 0))
```

Create splits.
```{r}
model_split <- h2o.splitFrame(model_h2o, destination_frames=c("train", "test"), seed=2121)
train <- model_split[[1]]
test <- model_split[[2]]
```

Load saved models. These do not include the channel flag.

```{r}
tree_model <- h2o.loadModel("/Users/jamespearce/repos/simu-late/GBM_grid_0_AutoML_20180914_074804_model_0")
```

Get predictions from the model.
```{r}
preds <- h2o.predict(tree_model, newdata=model_h2o)
model_h2o$expected_value <- preds
```
```{r}
y <- "response"
x <- c("Email", "Mail", "SMS", "TM")
channel_regression <- h2o.glm(x, y, model_h2o, model_id="channel_regression", seed=1438, offset_column="expected_value", intercept=FALSE, alpha=0, lambda=0)
```

```{r}
channel_regression
```
What happens without the offset term?
```{r}
channel_regression_0 <- h2o.glm(x, y, model_h2o, model_id="channel_regression", seed=1438, intercept=FALSE, alpha=0, lambda=0)
channel_regression_0
```


```{r}
h2o.shutdown()
```

