---
title: "Optimization Script 01"
output: html_notebook
---

Adjust the script from Bob Agnew to take in the dummy data developed for the IBM marketing optimization example. https://github.com/fzhang612/r_utility_code/blob/master/mkt_opt.R

> In the future may be able to adapt other constraints to this example, but need to understand the dual solution. In the meantime, can brute force it given the speed of computation.

The algorithm imbedded in this script optimizes the assignment of offers to
prospects, subject to stipulated offer quantity constraints and the limit of
at most one offer per prospect.

This script can be adapted for a different number of offers and alternative
equality or inequality constraints, as long as they can be linearized, although
the dual must be formulated for solution.  Inequality constraints would require
a different R solver (`nlminb`) which incorporates lower bounds. It is also possible to allow more than one offer per prospect.  


| Run | Size | Time (s)  | Value  | Dual value | Dual budget |
|-----|------|-----------|--------|------------|-------------|
|   1 |  27  |  0.9      |    603 |            |             |
|   2 |  50  |  1.8      |  1 283 |  1 168     |             |
|   3 |  75  |  0.25     |  1 975 |            |             |
|   4 |  85  |  0.55     |  2 227 |            |             |
|   5 |  90  |   --      |  --    |            |             |

Now that we have tested on $n = 50$, use the full 10,000 in the simulated file.

```{r setup}
knitr::opts_chunk$set(include = FALSE)
library(magrittr)
library(dplyr)
library(data.table)
library(dtplyr)
library(stringr)
```

```{r}
date()
```

Set up number of prospects.

```{r}
n_obs_new <- 10000
optimized_count <- 400 # the number of customers used in the explicit optimization

## see Python for more details
``` 

We have four product types:

  * car loan
  * savings
  * mortgage
  * pension
  
Each product has a different `productValue`: the revenue that can be obtained for the product on average. To get a fair representation of marketing across the various offers, each is allocated a `budgetShare`. 

```{r}
products <- c("Car loan", "Savings", "Mortgage", "Pension")
productValue <- c(100, 200, 300, 400)
budgetShare <- c(0.6, 0.1, 0.2, 0.1)
```

Each product these can be offered over one of the following channels:

  * gift
  * newsletter
  * seminar
  
Each of these channels has different costs, and each has a different _influence factor_. We use the influence to weight the estimated value of the response accordingly.

```{r}
channels <- c('gift', 'newsletter', 'seminar')
cost <- c(20, 15, 23)
factor <- c(0.2, 0.05, 0.3)
# channel_share <- rep(1/3, 3)
```

Now have 12 offers: product $\times$ channel.

```{r}
# Calculate the number of offers: channel x product
offer_scale <- n_obs_new/optimized_count
offers <- c(24, 4, 1, 11, 40, 0, 0, 0, 0, 79, 129, 53)*offer_scale %>%
  round 
```




Budget needs to be less than the available marketing budget of $500.

```{r}
availableBudget <- 500
```

Read in the offers data, originally from IBM and massaged. It gives the probability of taking an offer by each customer.

Rather than using the full 10,000, test that it works on a smaller size.

```{r}
product_probs_orig <- fread('offers_ibm_pivot.csv')
n_obs_original <- nrow(product_probs_orig)

product_probs <- fread('sample_data_10000.csv')

product_probs <- product_probs[seq(n_obs_new)]
n_obs <- nrow(product_probs)

adjustment_factor <- n_obs/n_obs_original
availableBudget <- availableBudget*adjustment_factor

# product_probs.rename(columns={'Unnamed: 0': 'customerid'}, inplace=True)
setnames(product_probs, "V1", "customerid")
head(product_probs)
```

Calculate the profit vectors.

```{r}
product_profit <- sweep(product_probs[, products, with = FALSE], 2, 
                        productValue, "*")
product_profit_1 <- product_profit*factor[1]
product_profit_2 <- product_profit*factor[2]
product_profit_3 <- product_profit*factor[3]
names(product_profit_1) <- paste(names(product_profit_1), channels[1])
names(product_profit_2) <- paste(names(product_profit_2), channels[2])
names(product_profit_3) <- paste(names(product_profit_3), channels[3])

product_profit <- cbind(product_profit_1,
                        product_profit_2,
                        product_profit_3)
```

### Dual Function

```{r}
dual <- function(u) {
  if (dim(product_profit)[2] != length(u)) {
    print(c(dim(product_profit)[2], length(u)))
    stop("Mismatched dimensions")
    }
  d <- sweep(product_profit, 2, u)
  v <- apply(d, 1, max) 
  v[v < 0] <- 0
  y <- offers%*%u + sum(v)
  y
}
```

```{r}
# Test dual
u_test <- c(11.2, 15, 6.02, 19.5, 0, 4.98, 2.23, 7.75, 50.2, 23.2, 9.09, 35.2)
dual(u_test)
```
```{r}
u_test <- rep(0, 12)
dual(u_test)
```
```{r}
apply(product_profit, 1, max) %>% head(10)
```
### Dual Minimization using Nonlinear Minimization Function

Here we are using the function `nlm` which is non-linear minimization. It takes a function to minimize and starting parameter values for the minimization as arguments.

```{r}
u_init <- offers*0
out <- nlm(dual, u_init, print.level = 1)
```

### Keep dual minimum and estimates.
```{r}
mindual <- out$minimum
u <- out$estimate
mindual
u
```

```{r}
d <- sweep(product_profit, 2, u)
d_DT <- data.table(d)
d_DT[, customerid := product_probs$customerid]
# d_DT[, v := v]

# d_DT_melt <- melt(d_DT, id.vars = c("customerid", "v"))
d_DT_melt <- melt(d_DT, id.vars = c("customerid"))
d_DT_alloc <- d_DT_melt[order(customerid, -value)][, lapply(.SD, head, 1), by = .(customerid)]

# check counts
d_DT_alloc[, .N, by = .(variable)]

# test
# use contrasts/one-hot encoding
d_DT_indicator <- model.matrix(~ variable - 1, d_DT_alloc) %>%
  as.data.table 
offer_names <- levels(d_DT_alloc$variable)
setnames(d_DT_indicator, offer_names)
d_DT_alloc_0 <- cbind(d_DT_alloc, d_DT_indicator)

setorder(d_DT_alloc_0, -value)

# Cumulatively sum each and compare with the target levels.
d_DT_alloc_0[, c(offer_names) := lapply(.SD, cumsum), 
             .SDcols=offer_names]

# Now I need to find which one hits the maximum offers first.
n_alloc <- nrow(d_DT_alloc_0)
offers_min <- offers*0 # copy shape and names
offers_min[] <- n_alloc # fill with maximum
for (j in seq_along(offers)){
  if (offers[j] > 0) {
    offers_min[j] <- min(seq(n_alloc)[d_DT_alloc_0[[j + 3L]] >= offers[j]])
  }
}
offers_min %>% which.min
offers_min %>% min
```
 This means we are hitting the limit with the pension seminar at position `r offers_min %>% min`. 
 
```{r}
# add an index to data table
d_DT_alloc_0[, idx := .I]
allocation_final_DT <- d_DT_alloc_0[idx <= min(offers_min)][, idx := NULL]

# update offers with remaining
# offers_rem <- offers - 
offers_alloc <- allocation_final_DT[, .N, by=.(variable)]
offers_rem <- offers
offers_match <- match(offers_alloc$variable, offer_names)
offers_rem[offers_match] <- offers_rem[offers_match] - offers_alloc$N
offers_rem

# could remove all the zeros earlier

```
 
Now repeat the exercise with the remainder. 

```{r}
repeat {
  d_DT_melt_rem <- d_DT_melt[!(customerid %in% allocation_final_DT[["customerid"]])]
  d_DT_melt_rem <- d_DT_melt_rem[!(variable %in% offer_names[offers_rem == 0])]
  
  d_DT_alloc <- d_DT_melt_rem[order(customerid, -value)][, lapply(.SD, head, 1), by = .(customerid)]
  
  d_DT_indicator <- model.matrix(~ variable - 1, d_DT_alloc) %>%
    as.data.table 
  offer_names <- levels(d_DT_alloc$variable)
  setnames(d_DT_indicator, offer_names)
  d_DT_alloc_0 <- cbind(d_DT_alloc, d_DT_indicator)
  
  setorder(d_DT_alloc_0, -value)

  # Cumulatively sum each and compare with the target levels.
  d_DT_alloc_0[, c(offer_names) := lapply(.SD, cumsum), 
               .SDcols=offer_names]
  
  n_alloc <- nrow(d_DT_alloc_0)
  offers_min <- offers*0 # copy shape and names
  offers_min[] <- n_alloc # fill with maximum
  for (j in seq_along(offers)){
    if (offers_rem[j] > 0) {
      offers_min[j] <- min(seq(n_alloc)[d_DT_alloc_0[[j + 3L]] >= offers_rem[j]])
    }
  }
  
  cat("Allocation reached for offer: ", offer_names[which.min(offers_min)], "\n")
  
  # add an index to data table
  d_DT_alloc_0[, idx := .I]
  allocation_final_DT <- rbind(allocation_final_DT, d_DT_alloc_0[idx <= min(offers_min)][, idx := NULL])
  
  # update offers with remaining
  # offers_rem <- offers - 
  offers_alloc <- allocation_final_DT[, .N, by=.(variable)]
  offers_rem <- offers
  offers_match <- match(offers_alloc$variable, offer_names)
  offers_rem[offers_match] <- offers_rem[offers_match] - offers_alloc$N
  # offers_rem
  
  if (sum(offers_rem) == 0) break
}
```




Calculate the cost and profit.

```{r}
library(stringr)

product_profit_DT <- data.table(product_profit)
product_profit_DT[, customerid := product_probs$customerid]
product_profit_DT <- melt(product_profit_DT, id.vars = c("customerid"))
setnames(product_profit_DT, "value", "profit")

product_profit_DT <- merge(allocation_final_DT, product_profit_DT, 
                           by = c("customerid", "variable"), 
                           all.x = TRUE)

channel_costs <- data.table(channels, cost)
setnames(channel_costs, c("Channel", "cost"))
# extract channel from variable

product_profit_DT[, Channel := word(variable, -1)]
product_profit_DT <- merge(product_profit_DT, channel_costs, 
                       by = "Channel", all.x = TRUE)
# setorder(product_profit_DT, -v)
product_profit_DT[, product := word(variable, 1,
                                    str_count(as.character(variable),
                                              boundary("word")) - 1)]
# product_profit_DT[, cum_cost := cumsum(cost)]
# product_profit_DT[, cum_profit := cumsum(profit)]
# product_profit_DT <- product_profit_DT[cum_cost <= availableBudget]


```

### Overall results
```{r}
product_profit_DT[, .(profit = sum(profit), budget = sum(cost), .N)]
```

### Results by channel
```{r}
product_profit_DT[, .(profit = sum(profit), 
                      budget = sum(cost), .N),
                  by=.(channel=Channel)
                  ]

```
### Results by product
```{r}
product_profit_DT[, .(profit = sum(profit), 
                      budget = sum(cost), .N),
                  by=.(product)
                  ]
```

