---
title: "Assigning students to courses"
author: "Dirk Schumacher"
date: "2016-12-16"
output: html_notebook
---


## Introduction

In this article we will look at [assignment problems](https://en.wikipedia.org/wiki/Generalized_assignment_problem). 

As a real world example problem we would like to match a group of students to a set of courses with the following constraints:

* Each course has a capacity
* Every student needs to be assigned to exactly one course. 
* All students have stated individual preferences on a scale from 1 to 3, where 3 is the most favorable course.

## The data

```{r}
n_obs_new <- 50
``` 

We have four product types:

  * car loan
  * savings
  * mortgage
  * pension
  
Each product has a different `productValue`: the revenue that can be obtained for the product on average. To get a fair representation of marketing across the various offers, each is allocated a `budgetShare`. 

```{r}
products <- c("Car loan", "Savings", "Mortgage", "Pension")
productValue <- c(100, 200, 300, 400)
# budgetShare <- c(0.6, 0.1, 0.2, 0.1)
```

Each product these can be offered over one of the following channels:

  * gift
  * newsletter
  * seminar
  
Each of these channels has different costs, and each has a different _influence factor_. We use the influence to weight the estimated value of the response accordingly.

```{r}
channels <- c('gift', 'newsletter', 'seminar')
cost <- c(20, 15, 23)
factor <- c(0.2, 0.05, 0.3)
# channel_share <- rep(1/3, 3)
```

Budget needs to be less than the available marketing budget of $500.

```{r}
availableBudget <- 500
```

Read in the offers data, originally from IBM and massaged. It gives the probability of taking an offer by each customer.

Rather than using the full 10,000, test that it works on a smaller size.

```{r}
library(data.table)

product_probs_orig <- fread('offers_ibm_pivot.csv')
n_obs_original <- nrow(product_probs_orig)

product_probs <- fread('sample_data_10000.csv')

product_probs <- product_probs[seq(n_obs_new)]
n_obs <- nrow(product_probs)

adjustment_factor <- n_obs/n_obs_original
availableBudget <- availableBudget*adjustment_factor

# product_probs.rename(columns={'Unnamed: 0': 'customerid'}, inplace=True)
setnames(product_probs, "V1", "customerid")
head(product_probs)
```

Set up the numbers of channels, products and customers.

```{r}
num_customers <- nrow(product_probs)
num_products <- length(products)
num_channels <- length(channels)
```

```{r}
# Create a function to make extracting easier
profit <- function(i, j, k){
  pr <- factor[k]*productValue[j]*product_probs[i, products[j], with = FALSE]
  as.integer(round(pr))
}

profit(3, 1, 1)
```

```{r}
library(ompr)
channel_min <- 5
product_min <- 8
model <- MIPModel() %>%
  
  # 1 if customer i is product j via channel k
  add_variable(x[i, j, k], 
               i = seq(num_customers), 
               j = seq(num_products), k = seq(num_channels), 
               type = "binary") %>%
  
  # maximize the revenue
  set_objective(sum_expr(x[i, j, k]*profit(i, j, k),
                         i = seq(num_customers),
                         j = seq(num_products), k = seq(num_channels) )) %>%
 
  # maximum of one offer per customer
  add_constraint(sum_expr(x[i, j, k], j = seq(num_products),
                          k = seq(num_channels)) <= 1,
                 i = seq(num_customers)) %>%
  
  # do not exceed the budget
  add_constraint(sum_expr(x[i, j, k]*cost[k], i = seq(num_customers),
               j = seq(num_products),
               k = seq(num_channels)) <= availableBudget) %>%
  
# product and channel minimum constraints
  
model
```

## Solve

```{r}
library(ompr.roi)
library(ROI.plugin.glpk)
result <- solve_model(model, with_ROI(solver = "glpk", verbose = TRUE))
```

We solved the problem with an objective value of `r objective_value(result)`.

```{r}
objective_value(result)
```


```{r}
matching <- result %>% 
  # get_solution(x[i, j, k]) %>%
  get_solution(x[i, j]) %>%
  filter(value > 0) #%>%
#   select(i, j, k) %>% 
#   rowwise() %>% 
#   mutate(weight = weight(as.numeric(i), as.numeric(j)), 
#          preferences = paste0(preferences(as.numeric(i)), collapse = ",")) %>% ungroup
```

```{r}
head(matching)
```


```{r}
matching %>% 
  group_by(weight) %>% 
  summarise(count = n())
```
### The previous stuff --- hive off rather than delete _for now_
We have $n$ students:
```{r}
n <- 40
```

And $m$ courses with equal capacity. The capacity can vary among courses though.
```{r}
m <- 4
capacity <- rep.int(11, m) # all have equal capacities
```

In addition, each student has three preferences. To model this we have a function that gives us three courses for each student. The first component has perference 1, second 2, and third 3:
```{r}
set.seed(1234)
preference_data <- lapply(seq_len(n), function(x) sample(seq_len(m), 3))
preferences <- function(student) preference_data[[student]]
```

```{r}
preferences(1)
```

The last component we need is a weight functions to make the model formulation easier.
This function gives us the preference weighting for a course and student pair.

```{r}
# the weight of a student choosing a course
# if the course is not among the preferences, the weight is -100000
weight <- function(student, course) {
  p <- which(as.numeric(course) == preferences(as.numeric(student)))
  as.integer(if (length(p) == 0) {
    -100000
  } else {
    p
  })
}
```

Some examples:
```{r}
weight(1, 3)
weight(1, 23) # this was not a choice by student 1, so we give it a big penalty
```

## The model

The idea is to introduce a binary variable $x_{i, j}$ that is $1$ if student $i$ is matched to course $j$. As an objective we will try to satisfy preferences according to their weight. So assigning a student to a course with preference 3 gives 3 points and so forth. The model assumes, that the total capacity of the courses is enough for all students.

Here it is in mathematical notation:

$$
\begin{equation*}
\begin{array}{ll@{}ll}
\text{max}  & \displaystyle\sum\limits_{i=1}^{n}\sum\limits_{j=1}^{m}weight_{i,j} \cdot x_{i, j} & &\\
\text{subject to}& \displaystyle\sum\limits_{i=1}^{n}   x_{i, j} \leq capacity_j,  & j=1 ,\ldots, m&\\
                 & \displaystyle\sum\limits_{j=1}^{m}   x_{i, j} = 1,  & i=1 ,\ldots, n&\\
                 &                                                x_{i,j} \in \{0,1\}, &i=1 ,\ldots, n, & j=1 ,\ldots, m
\end{array}
\end{equation*}
$$

Or directly in R:

```{r, message=FALSE}
library(ompr)
model <- MIPModel() %>%
  
  # 1 iff student i is assigned to course m
  add_variable(x[i, j], i = 1:n, j = 1:m, type = "binary") %>%
  
  # maximize the preferences
  set_objective(sum_expr(weight(i, j) * x[i, j], i = 1:n, j = 1:m)) %>%
  
  # we cannot exceed the capacity of a course
  add_constraint(sum_expr(x[i, j], i = 1:n) <= capacity[j], j = 1:m) %>% 
  
  # each student needs to be assigned to one course
  add_constraint(sum_expr(x[i, j], j = 1:m) == 1, i = 1:n)
model
```

## Solve the model

We will use `glpk` to solve the above model.

```{r}
library(ompr.roi)
library(ROI.plugin.glpk)
result <- solve_model(model, with_ROI(solver = "glpk", verbose = TRUE))
```

We solved the problem with an objective value of `r objective_value(result)`.

```{r}
matching <- result %>% 
  get_solution(x[i,j]) %>%
  filter(value > .9) %>%  
  select(i, j) %>% 
  rowwise() %>% 
  mutate(weight = weight(as.numeric(i), as.numeric(j)), 
         preferences = paste0(preferences(as.numeric(i)), collapse = ",")) %>% ungroup
```

```{r}
head(matching)
```


```{r}
matching %>% 
  group_by(weight) %>% 
  summarise(count = n())
```



`r nrow(filter(matching, weight == 3))` students got their top preference.
`r nrow(filter(matching, weight == 2))` students were assigned to their second choice and `r nrow(filter(matching, weight == 1))` students got their least preferable course.

The course assignment now looks like this:

```{r}
plot_data <- matching %>% 
  mutate(course = factor(j), weight = factor(weight, levels = c(1, 2, 3))) %>% 
  group_by(course, weight) %>% 
  summarise(count = n()) %>% 
  tidyr::complete(weight, fill = list(count = 0))
ggplot(plot_data, aes(x = course, y = count, fill = weight)) + 
  geom_bar(stat = "identity") + 
  viridis::scale_fill_viridis(discrete = TRUE) + 
  geom_hline(yintercept = 11)
```

