---
title: "Optimization Script 00"
output: html_notebook
---

I downloaded this script from https://github.com/fzhang612/r_utility_code/blob/master/mkt_opt.R


## MARKETING OPTIMIZATION R-SCRIPT

Thanks to Bob Agnew, raagnew1@gmail.com, www.raagnew.com.

The algorithm imbedded in this script optimizes the assignment of offers to
prospects, subject to stipulated offer quantity constraints and the limit of
at most one offer per prospect [_we would want to change this by parameter_].  Remarkably, this demo problem, with a million prospects and three offers, solves in less than a minute on my home PC.  Previously,
I solved it using SAS PROC NLP on a big server and the time required was much
greater. Ref:
www.raagnew.com/uploads/5/7/1/7/57174385/dual_solution_marketing_optimization.pdf

This script can be adapted for a different number of offers and alternative
equality or inequality constraints, as long as they can be linearized, although
the dual must be formulated for solution.  Inequality constraints would require
a different R solver (`nlminb`) which incorporates lower bounds. It is also possible to allow more than one offer per prospect.  

```{r}
date()
```

Set up number of prospects.

```{r}
n <- 1000000
``` 


Initialize prospect offer and profit at zero, i.e., no offer.

```{r}
Profit <- Offer <- rep(0,n)

```

### Simulate Prospect Profits for Three Offers

Different profit vectors will be generated for each execution, which has the
advantage of showing that solution is robust across various instances.

```{r}
p1 <- 100*runif(n,0,1) - 10
p2 <- .6*p1 + .4*(60*runif(n,0,1) - 6)
p3 <- .4*p1 + .6*(40*runif(n,0,1) - 4)

```

### Mean Offer Profits

```{r}
c(mean(p1),mean(p2),mean(p3))

```

### Stipulated Offer Quantities

```{r}
b <- c(300000,200000,100000)
```

### Dual Function

```{r}
dual <- function(u) {
  d1 <- p1 - u[1]
  d2 <- p2 - u[2]
  d3 <- p3 - u[3]
  v <- pmax(d1,d2,d3,rep(0,n))
  y <- b%*%u + sum(v)
  y
}
```

### Dual Minimization using Nonlinear Minimization Function

Here we are using the function `nlm` which is non-linear minimization. It takes a function to minimize and starting parameter values for the minimization as arguments.

```{r}
out <- nlm(dual,c(0,0,0), print.level = 1)
# out
```

### Keep dual minimum and estimates.
```{r}
mindual <- out$minimum
u <- out$estimate
d1 <- p1 - u[1]
d2 <- p2 - u[2]
d3 <- p3 - u[3]
v <- pmax(d1,d2,d3)
ord <- order(v, decreasing = TRUE)
s <- c(0,0,0)
h <- 100000000
for (j in 1:sum(b)) {
  k <- s < b
  m <- ord[j]
  i <- which.max(c(k[1]*d1[m]-(1-k[1])*h,k[2]*d2[m]-(1-k[2])*h,k[3]*d3[m]-(1-k[3])*h))
  Offer[m] <- i
  Profit[m] <- c(p1[m],p2[m],p3[m])[i]
  s[i] <- s[i] + 1
}
```


```{r}
mindual
# Primal Optimum (Maximum Profit)
sum(Profit)
# Note that we are extremely close, in some instances right on.
# Offer Quantities
s
# Offer Profits
c(sum(Profit[Offer == 1]), sum(Profit[Offer == 2]), sum(Profit[Offer == 3]))
```


